---
title: "Untitled"
output: html_document
---
#The New York Times web site provides a rich set of APIs, as described here: http://developer.nytimes.com/docs You'll need to start by signing up for an API key. Your task is to choose one of the New York Times APIs, construct an interface in R to read in the JSON data, and transform it to an R dataframe.
```{r}
library(tidyverse)
library(jsonlite)
library(knitr)
library(dplyr)
library(httr)
library(rjson)
library(stringi)
```
#Get my key from New York Times website
```{r}
my_key = "VyfZDT2zI2fvSp2OyKogKtzD6HrmTu5l"
```
#read json files of all movie reviews with my key
```{r}
df <- jsonlite::fromJSON(txt = paste0("https://api.nytimes.com/svc/movies/v2/reviews/all.json?api-key=", my_key)) %>% 
  data.frame

df %>%
  head()
```
#Retrieve column names on data frame
```{r}
colnames(df)
```
#The link list contains the link information to the movie review article
```{r}
kable(names(df$results.link), format="markdown")
```
#Get information about the number of movie reviews returned by API data request
```{r}
nrow(df)
```
#Display all the results from the Link list
```{r}
all <- cbind(df[1:9], df$results.link$type, df$results.link$url,df$results.link$suggested_link_text)
kable(head(all), format="markdown")
```
#Retrieve some information from the title, byline, headline, summary, date, and link
```{r}
friendly_display <- cbind(df[c(5,8,9,10,12)], df$results.link$url)
kable(friendly_display, format="markdown")
```










