---
title: "Project 2C - Energy Use by Sector"
author: "Jill Anderson"
date: "Due October 8, 2017"
output: html_document
---

## INTRODUCTION
#### Data: Table C1. Energy Consumption Overview: Estimates by Energy Source and End-Use Sector, 2015
#### Units are in Trillion Btu
#### US Energy Information Administration

## PREPARE
#### Load required lirbaries.
```{r}
library(knitr)
library(stringr)
library(tidyr)
library(dplyr)
library(ggplot2)
library(DT)
```

## LOAD DATA
#### Load raw data file into R
```{r}
energy_all <- read.csv("~/Desktop/EIA-source-use.csv", header = FALSE, stringsAsFactors = FALSE, check.names = FALSE)
energy_all[1:10,]
```

#### This shows the header is across three rows 
#### Read the file again without the top two header rows
```{r}
energy_raw <- read.csv("~/Desktop/EIA-source-use.csv", header = TRUE, skip = 2, stringsAsFactors = FALSE, check.names = FALSE)
energy_raw[1:10,]
```

## SUBSET DATA
#### Compare states by looking at the consumption of energy by the four types by removing the source columns
```{r}
energy_use <- energy_raw[,-2:-10]
energy_use[1:10,]
```

## TIDY DATA
#### Gather and tidy data around the different categories of usage
```{r}
energy_use_tidy <- gather(energy_use, "Use_Type", "Total_By_Type", Residential:Transportation) %>% 
  arrange(State)
datatable(energy_use_tidy)
```

## CALCULATE TOTAL
#### Calculate the total energy consumed for each state by adding every four rows
```{r}
total_state <- colSums(matrix(energy_use_tidy$Total_By_Type, nrow=4))
```

#### Create a vector with the total value repeated four times to add to the working data set
```{r}
total_state_4 <- rep(total_state, each = 4)
total_state_4 
energy_use_tidy$Total_All <- total_state_4 
datatable(energy_use_tidy)
```

## CALCULATE PERCENTAGE
#### Calculate the percentage contribution for each use type against the total
```{r}
energy_use_perc <- energy_use_tidy %>%
  mutate(Percent_Total = Total_By_Type / Total_All)
```

#### Round to shorten the Percentage of Total calculation to two digits
```{r}
energy_use_perc[,"Percent_Total"]=format(round(energy_use_perc[,"Percent_Total"]*100,2),nsmall=2)
datatable(energy_use_perc)
```

## ANALYZE DATA
#### Various analyses can be performed at this point, such as identifying the state that has the maximum percentage contribution from one of the four sectors 
```{r}
energy_use_perc$State[which.max(energy_use_perc[,"Percent_Total"])]
energy_use_perc$Use_Type[which.max(energy_use_perc[,"Percent_Total"])]
paste(max(energy_use_perc[,"Percent_Total"]),"%",sep="")
```

#### Similarily looking at which state has the minimum consumption percentage from any one sector
```{r}
energy_use_perc$State[which.min(energy_use_perc[,"Percent_Total"])]
energy_use_perc$Use_Type[which.min(energy_use_perc[,"Percent_Total"])]
paste(min(energy_use_perc[,"Percent_Total"]),"%",sep="")
```

#### Additionally, we can look at which state has the highest total consumption percentage from any one sector
```{r}
energy_use_perc$State[which.max(energy_use_perc[,"Total_By_Type"])]
energy_use_perc$Use_Type[which.max(energy_use_perc[,"Total_By_Type"])]
max(energy_use_perc[,"Total_By_Type"])
paste(energy_use_perc$Percent_Total[which.max(energy_use_perc[,"Total_By_Type"])],"%",sep="")
```

#### Additionally, we can look at which state has the highest total consumption from any one sector
```{r}
energy_use_perc$State[which.max(energy_use_perc[,"Total_By_Type"])]
energy_use_perc$Use_Type[which.max(energy_use_perc[,"Total_By_Type"])]
max(energy_use_perc[,"Total_By_Type"])
```
